Tools for reproducible ex-post survey data harmonization

### 2020-07-13 data-raw/article.Rmd
update article (just links and a bit of code description, nothing fundamental.)

* Creating a full [documentation](#documentation) of conversions, recoding for reproducability.

@@

The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html), [as_character - *documentation missing*], and [as_factor](http://retroharmonize.satellitereport.com/reference/as_factor.html) methods, by replacing the user-defined missing values with type-specific NA values in the R language (NA_real_ or NA_character_). After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.

@@

Working with [harmonize_values()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) is relatively easy if the researcher wants to join a few variables in a few survey files, but for large-scale, programmatic use it not really suitable.

@@

<just simple description and code links>

* exporting to SPSS
* exporting to Stata
* exporting to flat files

@@

Our simple methods, [as_factor()](http://retroharmonize.satellitereport.com/reference/as_factor.html) (imported from haven) and [as_numeric()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) precisely do this. While we would not advocate the use of character vectors for statistical analysis, for data visualization they may be more practical, and as_character() [*not het documented*] will display the value labels as a character vector.

>>>>

* Creating a full [documentation](#documentation) of conversions, recoding for reproducibility.

@@

The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods, by replacing the user-defined missing values with type-specific `NA` values in the R language (`NA_real_` or `NA_character_`). After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.

SPSS and the [labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) class allows the parallel use of two attributes, `na_values` for enumeration of all values that should be treated as missing, or `na_range` for declaring a continuous range as missing. In surveying, the use of `na_values` is more practical, nevertheless, we must check the consistency of these attributes when we want to bind together several vectors.  The [na_range_to_values](http://retroharmonize.satellitereport.com/reference/na_range_to_values.html) creates this consistency in case `na_range` is present in a vector. We did not implement the reverse `na_values` to `na_range` as it is unlikely to be used for survey harmonization. The function gives a warning in case of inconsistency --- in such a case we recommend a manual review of the valid and missing ranges of the vector, as this is more of a logical than a syntactical error in the survey file.

@@

Working with [harmonize_values()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) is relatively easy if the researcher wants to join a few variables in a few survey files, but for large-scale, programmatic use it not really suitable. (See [vignette](http://retroharmonize.satellitereport.com/articles/harmonize_labels.html))

@@

>Unfortunately, the C++ code that writes from haven to sav (behind write_sav) is not implemented for the haven_labelled_spss class and does not record back user-defined NA values.  I was not able to amend the code, but I wrote a detailed bug report.

>Currently, for testing and demonstration purposes, we must rely on the original GESIS files, or, if we want to show smaller or integrated examples, their rds representations.

See [read_rds](http://retroharmonize.satellitereport.com/reference/read_rds.html) for the concept, the *write_rds* pair is not yet implemented, but easy to imagine. My [read_spss](http://retroharmonize.satellitereport.com/reference/read_spss.html) does what I imagined, but I cannot modify the haven::write_sav in C++ so until it is fixed, we cannot rely on that.

> End of comments.

@@

Our simple methods, [as_factor()](http://retroharmonize.satellitereport.com/reference/as_factor.html) (imported from haven) and [as_numeric()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) precisely do this. While we would not advocate the use of character vectors for statistical analysis, for data visualization they may be more practical, and [as_character()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) will display the value labels as a character vector.
